---
title: "Phylogenetic Tree Representation"
author:
- name: Eric Marcon
  affiliation: AgroParisTech, UMR EcoFoG
  email: eric.marcon@ecofog.gf
package: entropart
output:
  BiocStyle::html_document
abstract: |
  Several object classes can be used in `r CRANpkg("entropart")` to represent phylogenies and calculate phylogenetic diversity. They are detailed here.
vignette: |
  %\VignetteIndexEntry{Phylogenies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Formats used

Supported phylogenies are classes `phylo` from package `r CRANpkg("ape")`, `phylog` from package `r CRANpkg("ade4")` and `hclust` from package `r Rpackage("base")`.

To optimize performance, a specific class `PPtree`, for "preprocessed trees", is introduced by the package: it is basically a list containing a `phylo` and an `hclust` representation of the phylogeny and preprocessed useful statistics.

The arguments `Tree` or `PhyloTree` used in phylogenetic diversity estimation functions of the package may be any of those formats.


# Conversion between classes

Let's start from an ultrametric distance matrix.
Three species are defined, with distance 1 between the first 2 and distance 2 between both and the last one.

```{r DistMatrix}
dSp <- matrix(c(0, 1, 2, 1, 0, 2, 2, 2, 0), nrow=3, byrow=TRUE)
row.names(dSp) <- colnames(dSp) <- c("A", "B", "C")
dSp
```

## hclust

An `hclust` object is created by UPGMA hierarchical clustering.

```{r hclust}
require("stats")
plot(hTree <- hclust(as.dist(dSp), method="average"), hang=-0.01, axes = F)
axis(2)
```

Node heights are stored in `$height`.

```{r hclust2}

hTree$height
```


## phylo

Conversion to a `phylo` object is straightforward.

```{r phylo}
require("ape")
plot(phyloTree <- as.phylo(hTree))
axis(1)
```

Edge lengths have been divided by 2 during the conversion, as documented in `?as.phylo.hclust`.
```{r phylo2}
phyloTree$edge.length
```

That does not suit our needs. In `r CRANpkg("entropart")`, edge lengths are multiplied by 2 after conversion, so that `phylo`objects can be identical to other phylogenies. 

```{r phylo3}
phyloTree$edge.length <- 2*phyloTree$edge.length
plot(phyloTree)
axis(1)
```

## phylog

The last conversion is from `phylo` to `phylog`.

```{r phylog}
require("ade4")
plot(phylogTree <- hclust2phylog(hTree))
axis(1)
```

Edge lengths are not stored directly in `phylog` objects.
The `$droot` component keeps distances from tips to root.
The `$Wdist` component is a `dist` (half a distance matrix without the zero-diagonal) whose values are $d_{s,t}=\sqrt{2 \delta_{s,t}^2}$, where $\delta_{s,t}$ is the distance between species $s$ ant $t$.

```{r phylog2}
phylogTree$droot
phylogTree$Wdist^2/2
```

# Newick dendrograms

A Newick description of a dendrogram can read into a `phylo` object.

```{r Newick}
NewickABC <- "(C:2,(A:1,B:1):1);"
plot(phyloABC <- ape::read.tree(text=NewickABC))
axis(1)
````
